---
title: "individual2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(maps)
```

Load the two datasets.

```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
pers <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

```

1.Are there some days of the week where more accidents happen than the others (see FARS manual, use variable DAY_WEEK)?


```{r}
table(acc$DAY_WEEK)



```
By looking at the table above, we can check out that most of the accidents tend to take place during Friday ~ Sunday, where we call those days weekends. 


2.Create a data frame containing the persons who are fatally hurt in the accidents (look up variable INJ_SEV)
```{r}
fat<- pers %>%
  filter(INJ_SEV == 4)

fat



```
By looking at the dataset, it seems like most of the people who got fatally hurt in the accidents seems to have drink alcohol before the incident. 



3.Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make.
Make sure to handle the missing values appropriately. (look up variable MAKE)

```{r}
danger_veh <- fat %>%
  drop_na(MAKE) %>%
  group_by(STATE, MAKE) %>%
  summarise(count = n()) %>%
  group_by(STATE) %>%
  filter(count == max(count))

danger_veh
```
From the data frame table above, we can see that some state codes are missing, which is 3,7,14,43,52. The state codes that are missing are "American Samoa," "Canal Zone," "Guam," "Puerto Rico," and "Virgin Islands of the US" as following. Since we are trying to look at the data in the inland USA, those values are not in the dataset. 

Not only the missing values for the STATE code but there are also duplicated values within the table. Those values are for the variable "MAKE," which stands out for the code for the branch that has the highest amount of fatal accidents. Since the number of counts is the same, we had decided to include both of the analysis branches. 




4.Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)

```{r}
ggplot(acc,aes(x=LONGITUD, y=LATITUDE))+
  geom_polygon(aes(group = STATE))

```



5.Join the accident and person table (work out which variable(s) to use)

```{r}

```



6. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results and explain what you find.

```{r}

```


7. Now plot a choropleth map of the number of deaths on a county level. Also explain what you find.

```{r}

```


8.Is summer or winter more dangerous? Does this depend on states? Explore and explain.

```{r}

```

