---
title: "individual1.rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
```

```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
pers <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```


Question 1
```{r}
table(acc$DAY_WEEK)
```

from the above data, we can see that most accidents occur on friday, saturday, and sunday


Question 2
```{r}
fatal <- pers %>%
  filter(INJ_SEV == 4)

fatal
```

Question 3
```{r}
dang_vehicle <- fatal %>%
  drop_na(MAKE) %>%
  group_by(STATE, MAKE) %>%
  summarise(count = n()) %>%
  group_by(STATE) %>%
  filter(count == max(count))

dang_vehicle
```

```{r}
glc <- readxl::read_xlsx("/Users/visma/desktop/ds202/ds202_lab5/FRPP_GLC_-_United_StATESaPRIL122021.xlsx")
```


Question 5
```{r}
accidents <-inner_join(pers, acc, by = c("STATE", "ST_CASE", "HOUR"))
accidents
```


Question 6
```{r}
tally1 <- accidents %>%
  group_by(DAY_WEEK) %>%
  summarize(DayCount = n())

ggplot(tally1, aes(x = DAY_WEEK, weight = DayCount, fill = DAY_WEEK)) + geom_bar()



tally2 <- accidents %>%
  group_by(HOUR) %>%
  summarize(HourCount = n())

ggplot(tally2, aes(x = HOUR, weight = HourCount, fill = HOUR)) + geom_bar()



tally3 <- accidents %>%
  group_by(SEX) %>%
  summarize(SexCount = n())

ggplot(tally3, aes(x = SEX, weight = SexCount, fill = SEX)) + geom_bar()
```

From the data given above, we can see that the most number of accidents usually occur in men around 7 PM on fridays, saturdays, and sundays



Question 7
```{r}
library(maps)
```


```{r}


death <- acc %>%
  group_by(STATE, COUNTY) %>%
  summarise(count = sum(FATALS))

county <- map_data("county")

glc2 <- glc %>%
  rename(STATE = `State Code`, COUNTY = `County Code`) %>%
  mutate(STATE = as.numeric(STATE), COUNTY = as.numeric(COUNTY))

DeathMap <- death %>%
  inner_join(glc2, by = c("STATE", "COUNTY")) %>%
  select(count, `State Name`, `County Name`) %>%
  distinct() %>%
  rename(region = `State Name`, subregion = `County Name`) %>%
  mutate(region = tolower(region), subregion = tolower(subregion)) %>%
  right_join(county, by = c("region", "subregion"))

DeathMap %>%
  ggplot(aes(x = long, y = lat, fill = count)) + geom_polygon(aes(group = group))
```


Question 8
```{r}
sumDeaths <- acc %>%
  filter(MONTH >= 6 | MONTH <= 8) %>%
  select(STATE, FATALS) %>%
  group_by(STATE) %>%
  summarise(sumCount = sum(FATALS))

winDeaths <- acc %>%
  filter(MONTH >= 11 | MONTH <= 2) %>%
  select(STATE, FATALS) %>%
  group_by(STATE) %>%
  summarise(winCount = sum(FATALS))

sumDeaths
winDeaths

totSumDeaths <- sum(sumDeaths$sumCount)
totSumDeaths

totWinDeaths <- sum(winDeaths$winCount)
totWinDeaths
```

there are significantly more deaths in the summer season than in winter, so this leads me to believe that summer is a more dangerous season to drive in


